An excess of positive results: Comparing the standard Psychology literature with Registered Reports

### 2020-07-12 submission2/positive_results_SRs_RRs_submission.Rmd
final minor edits

keywords          : "Publication bias, Registered Reports, hypothesis testing"
#wordcount         : "6488"

@@

Registered Reports (RRs) are a publication format with a restructured submission timeline: 
Before collecting data, authors submit a study protocol containing their hypotheses, planned methods, and analysis pipeline, which undergoes peer review.
If successful, the journal commits to publishing the final article following data collection, regardless of whether the hypotheses are supported ("in-principle acceptance").
The authors then collect and analyse the data and complete the final report. 
The final report is peer-reviewed again, but this time only to ensure that the the registered plan was adhered to and stated conclusions are justified (and, if applicable, that the data pass pre-specified quality checks).
Registered Reports thus combine an antidote to QRPs (preregistration) with an antidote to publication bias, because studies are selected for publication before their results are known. 
Since its introduction in 2013, the format has rapidly gained popularity and is offered by 250 journals at the time of writing (http://cos.io/rr).

@@

We set out to compare all published Registered Reports in Psychology with a new sample of standard reports obtained by replicating @Fanelli2010. 
Fanelli searched for articles containing the phrase "test$^\ast$ the hypothes$^\ast$", drew a random sample of 150 articles per discipline, and coded if the first hypothesis in each article had been supported or not.
For standard reports we used the same sampling method (restricted to the Psychology discipline), for Registered Reports we relied on a database curated by the Center for Open Science (COS). 
We chose this method because Fanelli's 2010 and 2012 studies (both use the same coding method) have been highly influential, and because it can easily be applied to a large set of studies. 
Because we expected many more Registered Reports than standard reports to be close replications of earlier studies$\,$---$\,$and perhaps movitated by scepticism of the original results$\,$---$\,$we additionally examined the role of replications in our analysis.

@@

# Methods
After conducting a pilot to test the planned procedure, we preregistered our study (https://osf.io/s8e97). 
Methods and analyses described here were preregistered unless otherwise noted. 
Our online materials include an Appendix with fine-grained methodological details and an annotated preregistration document with detailed comparisons to the eventual procedure (https://osf.io/dbhgr).

@@

### Additional measures
All additional measures we collected but have not described thus far were either auxiliary variables to facilitate the coding process or earlier versions of the variables discussed above, and are documented in the Appendix and in our shared dataset and codebook.

@@

`r printnum(n.hypintros.fivewords.RR, numerals = FALSE, capitalize = TRUE)` of the `r n.RR` Registered Reports ($`r printnum(n.hypintros.fivewords.RR/n.RR*100)` \%$) had at least one of these five most frequent word stems in their title, abstract, or keywords, meaning that a regular literature search (without access to full texts) with the search terms "hypothes$^\ast$ *OR* replicat$^\ast$ *OR* test$^\ast$ *OR* examine$^\ast$ *OR* predict$^\ast$" would have been effective in identifying these papers.
We do not know how well these search terms represent the population of hypothesis-testing studies in Psychology, but a structured investigation of this question could be useful for future meta-research.

@@

Another limitation of the current study [and of @Fanelli2010] is that standard reports were selected using the search phrase "test$^\ast$ the hypothes$^\ast$".
This phrase was virtually absent in  Registered Reports, suggesting that this search strategy may not yield a representative sample of the population of hypothesis-testing studies in the literature.
The use of the phrase might even be confounded with the outcome of a study:
For example, authors may be more likely to describe their research explicitly as a hypothesis test when they found positive results, but prefer more vague language for unsupported hypotheses (e.g., "we examined the role of ..."). 
A similar concern could be raised for the decision to code only the first reported hypothesis of each article.
The first hypothesis test may not be representative for all hypothesis tests reported in a paper, and the order of reporting may differ between standard reports and Registered Reports. 
For example, standard-report authors might tend to present supported hypotheses first, whereas Registered-Report authors might be more likely to present their hypotheses in "chronological" order.

>>>>

keywords          : "Publication bias, Registered Reports, hypothesis testing"
#wordcount         : "5870"

@@

Registered Reports (RRs) are a publication format with a restructured submission timeline: 
Before collecting data, authors submit a study protocol containing their hypotheses, planned methods, and analysis pipeline, which undergoes peer review.
If successful, the journal commits to publishing the final article following data collection, regardless of whether the hypotheses are supported ("in-principle acceptance").
The authors then collect and analyse the data and complete the final report. 
The final report is peer-reviewed again, but this time only to ensure that the the registered plan was adhered to and stated conclusions are justified (and, if applicable, that the data pass pre-specified quality checks).
Registered Reports thus combine an antidote to QRPs (preregistration) with an antidote to publication bias, because studies are selected for publication before their results are known. 
Since its introduction in 2013, the format has rapidly gained popularity and is offered by 256 journals at the time of writing (http://cos.io/rr).

@@

We set out to compare all published Registered Reports in Psychology with a new sample of standard reports obtained by replicating @Fanelli2010. 
Fanelli searched for articles containing the phrase "test$^\ast$ the hypothes$^\ast$", drew a random sample of 150 articles per discipline, and coded if the first hypothesis in each article had been supported or not.
For standard reports we used the same sampling method (restricted to the Psychology discipline), for Registered Reports we relied on a database curated by the Center for Open Science (COS). 
We chose this method because Fanelli's 2010 and 2012 studies (both use the same coding method) have been highly influential, and because it can easily be applied to a large set of studies. 
Because we expected many more Registered Reports than standard reports to be close replications of earlier studies$\,$---$\,$and perhaps motivated by scepticism of the original results$\,$---$\,$we additionally examined the role of replications in our analysis.

@@

# Methods
After conducting a pilot to test the planned procedure, we preregistered our study (https://osf.io/s8e97). 
Methods and analyses described here were preregistered unless otherwise noted. 
Our online materials include an Appendix with fine-grained methodological details and an annotated preregistration document with detailed comparisons to the eventual procedure (https://osf.io/dbhgr).
Appendix and open dataset also list all measures we collected but do not describe here (all of which were either auxiliary variables to facilitate the coding process or earlier versions of the variables discussed here).

@@

@@

`r n.hypintros.fivewords.RR` of the `r n.RR` Registered Reports ($`r printnum(n.hypintros.fivewords.RR/n.RR*100)` \%$) had at least one of these five most frequent word stems in their title, abstract, or keywords, meaning that a regular literature search (without access to full texts) with the search terms "hypothes$^\ast$ *OR* replicat$^\ast$ *OR* test$^\ast$ *OR* examine$^\ast$ *OR* predict$^\ast$" would have been effective in identifying these papers.
We do not know how well these search terms represent the population of hypothesis-testing studies in Psychology, but a structured investigation of this question could be useful for future meta-research.

@@

Another limitation of the current study [and of @Fanelli2010] is that standard reports were selected using the search phrase "test$^\ast$ the hypothes$^\ast$".
This phrase was virtually absent in  Registered Reports, suggesting that the search strategy may not yield a representative sample of the population of hypothesis-testing studies in the literature.
The use of the phrase might even be confounded with the outcome of a study:
For example, authors may be more likely to describe their research explicitly as a hypothesis test when they found positive results, but prefer more vague language for unsupported hypotheses (e.g., "we examined the role of ..."). 
A similar concern could be raised for the decision to code only the first reported hypothesis of each article.
The first hypothesis test may not be representative for all hypothesis tests reported in a paper, and the order of reporting may differ between standard reports and Registered Reports. 
For example, standard-report authors might tend to present supported hypotheses first, whereas Registered-Report authors might be more likely to present their hypotheses in "chronological" order.

### 2020-07-12 submission2/positive_results_SRs_RRs_submission.tex
final minor edits

Registered Reports (RRs) are a publication format with a restructured submission timeline:
Before collecting data, authors submit a study protocol containing their hypotheses, planned methods, and analysis pipeline, which undergoes peer review.
If successful, the journal commits to publishing the final article following data collection, regardless of whether the hypotheses are supported (\enquote{in-principle acceptance}).
The authors then collect and analyse the data and complete the final report.
The final report is peer-reviewed again, but this time only to ensure that the the registered plan was adhered to and stated conclusions are justified (and, if applicable, that the data pass pre-specified quality checks).
Registered Reports thus combine an antidote to QRPs (preregistration) with an antidote to publication bias, because studies are selected for publication before their results are known.
Since its introduction in 2013, the format has rapidly gained popularity and is offered by 250 journals at the time of writing (\url{http://cos.io/rr}).

@@

We set out to compare all published Registered Reports in Psychology with a new sample of standard reports obtained by replicating Fanelli (2010).
Fanelli searched for articles containing the phrase \enquote{test\(^\ast\) the hypothes\(^\ast\)}, drew a random sample of 150 articles per discipline, and coded if the first hypothesis in each article had been supported or not.
For standard reports we used the same sampling method (restricted to the Psychology discipline), for Registered Reports we relied on a database curated by the Center for Open Science (COS).
We chose this method because Fanelli's 2010 and 2012 studies (both use the same coding method) have been highly influential, and because it can easily be applied to a large set of studies.
Because we expected many more Registered Reports than standard reports to be close replications of earlier studies\(\,\)---\(\,\)and perhaps movitated by scepticism of the original results\(\,\)---\(\,\)we additionally examined the role of replications in our analysis.

@@

After conducting a pilot to test the planned procedure, we preregistered our study (\url{https://osf.io/s8e97}).
Methods and analyses described here were preregistered unless otherwise noted.
Our online materials include an Appendix with fine-grained methodological details and an annotated preregistration document with detailed comparisons to the eventual procedure (\url{https://osf.io/dbhgr}).

@@

\hypertarget{additional-measures}{%
\subsubsection{Additional measures}\label{additional-measures}}

All additional measures we collected but have not described thus far were either auxiliary variables to facilitate the coding process or earlier versions of the variables discussed above, and are documented in the Appendix and in our shared dataset and codebook.

@@

Sixty-nine of the 71 Registered Reports (\(97.18 \%\)) had at least one of these five most frequent word stems in their title, abstract, or keywords, meaning that a regular literature search (without access to full texts) with the search terms \enquote{hypothes\(^\ast\) \emph{OR} replicat\(^\ast\) \emph{OR} test\(^\ast\) \emph{OR} examine\(^\ast\) \emph{OR} predict\(^\ast\)} would have been effective in identifying these papers.
We do not know how well these search terms represent the population of hypothesis-testing studies in Psychology, but a structured investigation of this question could be useful for future meta-research.

@@

Another limitation of the current study (and of Fanelli, 2010) is that standard reports were selected using the search phrase \enquote{test\(^\ast\) the hypothes\(^\ast\)}.
This phrase was virtually absent in Registered Reports, suggesting that this search strategy may not yield a representative sample of the population of hypothesis-testing studies in the literature.
The use of the phrase might even be confounded with the outcome of a study:
For example, authors may be more likely to describe their research explicitly as a hypothesis test when they found positive results, but prefer more vague language for unsupported hypotheses (e.g., \enquote{we examined the role of \ldots{}}).
A similar concern could be raised for the decision to code only the first reported hypothesis of each article.
The first hypothesis test may not be representative for all hypothesis tests reported in a paper, and the order of reporting may differ between standard reports and Registered Reports.
For example, standard-report authors might tend to present supported hypotheses first, whereas Registered-Report authors might be more likely to present their hypotheses in \enquote{chronological} order.

>>>>

Registered Reports (RRs) are a publication format with a restructured submission timeline:
Before collecting data, authors submit a study protocol containing their hypotheses, planned methods, and analysis pipeline, which undergoes peer review.
If successful, the journal commits to publishing the final article following data collection, regardless of whether the hypotheses are supported (\enquote{in-principle acceptance}).
The authors then collect and analyse the data and complete the final report.
The final report is peer-reviewed again, but this time only to ensure that the the registered plan was adhered to and stated conclusions are justified (and, if applicable, that the data pass pre-specified quality checks).
Registered Reports thus combine an antidote to QRPs (preregistration) with an antidote to publication bias, because studies are selected for publication before their results are known.
Since its introduction in 2013, the format has rapidly gained popularity and is offered by 256 journals at the time of writing (\url{http://cos.io/rr}).

@@

We set out to compare all published Registered Reports in Psychology with a new sample of standard reports obtained by replicating Fanelli (2010).
Fanelli searched for articles containing the phrase \enquote{test\(^\ast\) the hypothes\(^\ast\)}, drew a random sample of 150 articles per discipline, and coded if the first hypothesis in each article had been supported or not.
For standard reports we used the same sampling method (restricted to the Psychology discipline), for Registered Reports we relied on a database curated by the Center for Open Science (COS).
We chose this method because Fanelli's 2010 and 2012 studies (both use the same coding method) have been highly influential, and because it can easily be applied to a large set of studies.
Because we expected many more Registered Reports than standard reports to be close replications of earlier studies\(\,\)---\(\,\)and perhaps motivated by scepticism of the original results\(\,\)---\(\,\)we additionally examined the role of replications in our analysis.

@@

After conducting a pilot to test the planned procedure, we preregistered our study (\url{https://osf.io/s8e97}).
Methods and analyses described here were preregistered unless otherwise noted.
Our online materials include an Appendix with fine-grained methodological details and an annotated preregistration document with detailed comparisons to the eventual procedure (\url{https://osf.io/dbhgr}).
Appendix and open dataset also list all measures we collected but do not describe here (all of which were either auxiliary variables to facilitate the coding process or earlier versions of the variables discussed here).

@@

@@

69 of the 71 Registered Reports (\(97.18 \%\)) had at least one of these five most frequent word stems in their title, abstract, or keywords, meaning that a regular literature search (without access to full texts) with the search terms \enquote{hypothes\(^\ast\) \emph{OR} replicat\(^\ast\) \emph{OR} test\(^\ast\) \emph{OR} examine\(^\ast\) \emph{OR} predict\(^\ast\)} would have been effective in identifying these papers.
We do not know how well these search terms represent the population of hypothesis-testing studies in Psychology, but a structured investigation of this question could be useful for future meta-research.

@@

Another limitation of the current study (and of Fanelli, 2010) is that standard reports were selected using the search phrase \enquote{test\(^\ast\) the hypothes\(^\ast\)}.
This phrase was virtually absent in Registered Reports, suggesting that the search strategy may not yield a representative sample of the population of hypothesis-testing studies in the literature.
The use of the phrase might even be confounded with the outcome of a study:
For example, authors may be more likely to describe their research explicitly as a hypothesis test when they found positive results, but prefer more vague language for unsupported hypotheses (e.g., \enquote{we examined the role of \ldots{}}).
A similar concern could be raised for the decision to code only the first reported hypothesis of each article.
The first hypothesis test may not be representative for all hypothesis tests reported in a paper, and the order of reporting may differ between standard reports and Registered Reports.
For example, standard-report authors might tend to present supported hypotheses first, whereas Registered-Report authors might be more likely to present their hypotheses in \enquote{chronological} order.
